Rates and delay times of type Ia supernovae in the Dark Energy Survey

### 2021-03-11 draft.tex
Updates from Overleaf

% Abstract of the paper
\begin{abstract}
We use a sample of 881 photometrically classified type Ia supernovae (SNe Ia) discovered by the Dark Energy Survey (DES) in the redshift range $0.2 < z <0.6$, along with 40415 field galaxies to calculate the rate of SNe Ia per galaxy. We recover the established positive trend of SN Ia rate as a function of galaxy stellar mass across a broad range of scales $8.5 \leq \log(M_*/\mathrm{M}_{\odot}) \leq 11.25$. We find that the SN Ia rate increases with stellar mass roughly linearly in log space, with a slope of $0.63 \pm 0.02$, which is consistent with previous work. We use an empirical model of stellar mass assembly to estimate the average star-formation histories (SFHs) of galaxies across the stellar mass range of our measurement. Combining the modelled SFHs with the SN Ia rates to estimate constraints on the SN Ia delay time distribution (DTD), we find the data are fit well by a power-law DTD with index $\beta = -1.14 \pm 0.05$ and normalisation $A = 2.1 \pm0.05 \times 10^{-13}~\mathrm{SNe}~{\mathrm{M}_{\odot}}^{-1}~\mathrm{yr}^{-1}$, which correspond to an overall SN Ia production efficiency $N_{\mathrm{Ia}}/M_* = 0.9~_{-0.7}^{+4.0} \times 10^{-3}~\mathrm{SNe}~\mathrm{M}_{\odot}^{-1}$. Upon splitting the SN sample by properties of the light curves, we find a strong dependence on DTD slope with the SN stretch parameter, with slower-declining SNe exhibiting a steeper DTD slope. We interpret this as a result of a relationship between stretch and progenitor age, and explore the implications of the result in the context of SN Ia progenitors.
\end{abstract}

@@

Type Ia supernovae (SNe Ia) are explosions of white dwarf stars (WDs). Although SNe Ia show diversity in their observed properties, a large fraction of them display a small dispersion in their peak brightnesses which can be reduced further through empirical relations between brightness and light curve properties, such as decline rate (stretch) or optical colour \citep{Rust1974,Pskovskii1977,Phillips1993,Tripp1998}. These properties have led SNe Ia to be used extensively by cosmologists to measure relative distances in the Universe \citep{Riess1998,Perlmutter1999}.

@@

The progenitors of a number of core-collapse SNe (CCSNe) have been identified in high-resolution pre-explosion images \citep{Smartt2009,Eldridge2013}, allowing detailed analysis of the latter stages of massive star evolution. However, this approach has not yet been successful in identifying a progenitor of a SN Ia (e.g. \citealt{Graur2014a,Kelly2014,Graur2019}, although see \citet{McCully2014} for evidence of a progenitor for a type Iax SN). Conversely, it has been possible to search for surviving remnants of the binary system -- main sequence stars or WDs left behind at the SN location or kicked out at high velocity \citep[e.g.][]{Schaefer2012,Ruiz-Lapuente2018,Kerzendorf2018,Kerzendorf2019}. \citet{Shen2018} discovered three high-velocity WDs consistent with models of double-degenerate, double-detonation SNe Ia. However, without an unambiguous observation of a SN Ia progenitor system or remnant companion, there is as yet no direct evidence ruling out any progenitor channel from providing a significant contribution to the total population of SNe Ia.

Another method of inferring progenitor channels is the study of the SN Ia delay time distribution (DTD). The DTD describes the rate at which SNe Ia occur as a function of the delay time $\tau$ since an episode of star-formation, and thus carries characteristic signatures of the progenitor channel (see \citealt{Wang2012} for a review of the theory, and \citet{Maoz2017} for a review of observations). The SD scenario produces a broad range of functional forms for the DTD, most of which fail to account for long delay times \citep{Graur2014}. On the other hand, most variants of the DD scenario predict a power-law of the form $\tau^{\beta}$, with $\beta \sim -1$ \citep[e.g][]{Ruiter2009,Mennekens2010}.

@@

\citet[][hereafter C14]{Childress2014} showed that the $A+B$ approximation arises as a direct consequence of the combination of a power-law DTD with the average SFHs of galaxies -- the prompt component proportional to the amount of on-going star-formation at the epoch of observation, and the tardy component caused by the fact that massive galaxies experienced high SFRs several Gyrs ago. The recent advances of integral field spectroscopy (IFS) allows for an extension of the SFHR method, by reconstructing the SFH for hundreds or thousands of local regions of each SN Ia host galaxy. Using this method, \citet{Castrillo2020} find a power-law slope of $-1.1\pm0.3$, while \citet{Chen2021} find $-1.4\pm0.3$.

The majority of observational evidence based on studies of SN Ia populations thus points towards a DD scenario for most, if not all, SNe Ia. However, finding self-consistent progenitor and explosion models that recreate the observed luminosity function as well as correlations between luminosity, light curve parameters, and host galaxy properties has proven difficult. In particular, simulations based around explosions of $M_{\mathrm{Ch}}$ WDs (linked strongly with the SD scenario but also with many DD scenarios) find difficulty in reproducing the light curves of "normal" SNe Ia as well as "peculiar" objects (\citealt{Ropke2007,Sim2013,Blondin2017}; see \citealt{Maoz2014,Jha2019} for overviews). In recent years, attention directed towards explosions of sub-$M_{\mathrm{Ch}}$ WDs triggered by double detonations (primarily related to a DD scenario) has led to promising results \citep[e.g.][]{Shen2017,Shen2018,Townsley2019,Gronow2020,Shen2021} although they still struggle to match observations at late times in the light curve evolution \citep{Shen2021,Gronow2021}. An additional factor in support of the double-detonation model is that the SN luminosity is related to the mass of the primary WD, which itself is likely to be related to its age (although this relation is probably complicated by other factors such as accretion rate, metallicity, and the composition of the companion), thereby providing an explanation for observed relation between light curve stretch and stellar age \citep{Rigault2013,Rigault2018,Rose2019,Nicolas2020}.

@@

In Section \ref{sec:data} we introduce our large sample of SNe Ia from the Dark Energy Survey (DES) as well as our deep sample of field galaxies that provide an effectively complete control sample from which to measure the rate. We describe our detailed handling of SN and galaxy incompleteness in Section \ref{sec:incompleteness}. We present the SN Ia rate per galaxy in Section \ref{sec:rates} and show that it is consistent with previous works. We outline our modelling of star-formation histories and our novel measurement of the power-law DTD in Section \ref{sec:model}. In Section \ref{sec:split_x1_c} we investigate how the DTD differs among sub-populations of SNe with different light curve characteristics, in particular the stretch, and show stretch to be strongly dependent on progenitor age. We conclude in Section \ref{sec:conclusion} by discussing the implications of the results.
Where relevant, we assume a spatially flat $\Lambda$CDM cosmology with $\Omega_m = 0.3$ and $H_0 = 70 \mathrm{ km s}^{-1}\mathrm{Mpc}^{-1}$. Unless otherwise stated, we assume Gaussian measurement uncertainties quoted at the $1\sigma$ level, and we quote the posterior median and $1\sigma$ credible intervals on derived parameters. Magnitudes are quoted in the AB system \citep{Oke1983}.

\begin{table}
	\centering
	\caption{Numbers of field galaxies and SN hosts passing various quality cuts. SNe are derived from must already have passed the masked chips and good redshift cuts before reaching this stage.}
	\label{tab:cuts}
	\begin{tabular}{lccr} % four columns, alignment for each
		\hline
		Cut & N (field galaxies)  & N (SN hosts)\\
		\hline
		Chosen fields & 1364311  & 1441\\
	    Kron mag in all bands & 1069004  & 1439 \\
	    Masked chips & 816950  & - \\
	    Has ugrizJHK photometry & 545748  & -\\
	    Edge of chip & 481731 & 1401 \\
	    Star/galaxy separation & 400051 &  1259\\
	    Has good redshift & 395034 & - \\
	    SNR $\geq 3$& 338256  & 1259 \\
	    $m_r \leq 24.5$ & 196109 &  1254 \\
	    $0.2 \leq z \leq 0.6$ & 48177 & 881\\

@@

In order to measure the rate of SNe Ia per galaxy per year, we require a sample of SNe Ia and a sample of field galaxies that are normalised to the same selection criteria -- the final rate calculation must be corrected for differences in sky area, redshift depth, and survey time. In this section, we introduce our SN and field galaxy samples.

\label{sec:data}
\subsection{Dark Energy Survey supernova programme \label{subsec:des}}
The Dark Energy Survey Supernova Programme (DES-SN) is a transient survey based on five six-month seasons of observations with the Dark Energy Camera (DECam; \citealt{Flaugher2015}), designed primarily to measure the light curves of SNe Ia for use as cosmological distance indicators. Transients were detected and processed using a difference imaging program \citep{Kessler2015} and vetted using a machine learning (ML) algorithm \citep{Goldstein2015}. Spectral follow-up of live SN candidates was performed on a suite of large optical telescopes \citep{Smith2020a}, leading to the initial publication of a measurement of cosmological parameters using 207 spectroscopically confirmed SNe Ia (\citealt{DESCollaboration2018a}, and references therein). Spectroscopic redshifts for host galaxies come from the OzDES Global Redshift Catalog (GRC; \footnote{\url{https://docs.datacentral.org.au/ozdes/overview/dr2/}} \citealt{Yuan2015,Childress2017,Lidman2020}), which comprises galaxies targetted as DES-SN hosts with the OzDES programme as well as redshifts from legacy catalogues in the DES-SN fields.
\subsection{Supernovae \label{subsec:host_sample}}
\subsubsection{Photometric classification and quality cuts \label{subsubsec:sn_classify}}
With over 30,000 transients passing ML selection it was not possible to obtain spectral follow up of every object. Instead we make use of photometric classification in the form of the recurrent neural network classifier \texttt{superNNova} (SNN; \citealt{Moller2019}), foloowing the approach of \citet{Scolnic2020}. A full description of the training and fitting of SNN will be presented in Vincenzi et al. \textit{in prep} and M\"{o}ller et al. \textit{in prep}; we provide a brief overview here.

@@

Since the vast majority of objects detected in the SN Deep catalogue do not have a spectroscopic redshift measurement, we rely on photometric redshifts (photo-$z$s).
Photo-$z$s are taken, from the DES Y3A2\_DEEP catalogue of \citet{Hartley2020}. Y3A2\_DEEP makes use of the same deep DES optical photometry that was used in SN Deep for a subset of the fields (SN-X3, SN-C3 and SN-E2), but adds DECam $u$-band data and near-infrared (NIR) $J$, $H$, and $K$-band imaging from the ultraVISTA and VIDEO surveys. The deep optical photometry in Y3A2\_DEEP was stacked using a different technique to that of SN Deep, but the resulting photometry is consistent within $1\sigma$ uncertainties \citep{Wiseman2020}. \citet{Hartley2020} estimated photo-$z$'s using the \texttt{eaZy} code \citep{Brammer2008}. Photo-$z$ accuracy from \citet{Hartley2020} at $17 < i<24$ is estimated around 0.03 as quantified by the Normalised Absolute Median Deviation (NMAD). This accuracy is degraded to 0.07 at $17 < i<26$, a range that includes all galaxies in our sample, but is more than adequate for the purposes of this work, where all field galaxies will be grouped into stellar mass bins of $\log \left(\Delta M_*\right) = 0.25$.

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/mass_hist_linlog_sn_field_BC03.png}
    \caption{Upper: Distribution of the stellar mass of the SN host (green dot-dashed) and field galaxies (orange solid) used in the analysis. Lower: As upper for field galaxies, but shown on a log scale.  The distribution closely matches data from the ZFOURGE survey at $0.2<z<0.5$ (cyan dashes) and $0.5<z<0.75$ (magenta dots).}
    \label{fig:mass_hist_hosts_field}
\end{figure}

@@

    \item field galaxies are limited to unmasked region of the SN-X3 chip \citep{Hartley2020};

@@

    \item to minimise stellar contamination, objects must have a value of 0.95 in the star/galaxy (S/G) separation \texttt{CLASS\_STAR} provided by \texttt{Source Extractor} \citep{Bertin1996};

@@

    \item objects must have a spectroscopic redshift measurement or a converged photometric redshift estimate;

@@

The numbers of SN hosts and field galaxies passing these cuts are listed in Table \ref{tab:cuts}. The final samples comprise 881 SNe and their host galaxies and 40,415 field galaxies. The volume-weighted mean redshifts are 0.50 for both SNe and field galaxies.

@@

We estimate global galaxy properties for both the SN host and field galaxies that pass the quality cuts by fitting the photometry with stellar population templates in a method outlined by \citet{Sullivan2006} and consistent to that used in previous DES-SN analyses \citep{Kelsey2021,Smith2020,Wiseman2020}. For SN hosts the redshift is fixed at the spectroscopically determined value, while for field galaxies we fix it at either the spectroscopic value if one exists in the OzDES GRC, or more commonly the photometrically derived value. We use the stellar population templates of \citet{Bruzual2003} and adopt a \citet{Chabrier2003} initial mass function (IMF). The fitting procedure returns a best fitting template and corresponding stellar mass ($M_*$). Upper and lower bounds on stellar mass estimates are taken as the extrema in the model parameters that are consistent with the data (given the photometric uncertainties) according to their $\chi^2$ statistic, as per \citet{Sullivan2006}. To check that the choice of templates does not bias the result, we also fit the galaxies using the P\'EGASE.2 templates \citep{Fioc1997,LeBorgne2002} and a \citet{Kroupa2001} IMF, and find results consistent within measurement uncertainties. This is consistent with the findings of \citet{Smith2020}.

@@

\section{Incompleteness corrections}
\label{sec:incompleteness}
The simple ratio of the number of SN hosts and field galaxies presented at the end of the previous section provides a first approximation of the SN rate per galaxy, and the addition of a factor equal to the survey duration normalises the rate to per year. However, both the SN and field galaxy samples introduced in Sections \ref{subsec:host_sample} and \ref{subsec:field_sample} are affected by incompleteness, which is likely to be the dominant systematic effect in the analysis. In this section we describe our method of correcting for various sources incompleteness in the data.

@@

To correct for these incompleteness, we follow a similar method to that used in the Palomar Transient Factory rates analyses of \citet{Frohmaier2019} and \citet{Frohmaier2020}. We run a suite of simulations using an identical method to that described in detail in \citet{Vincenzi2020}. We simulate SNe using the SuperNova ANAlysis software (\texttt{SNANA} \citealt{Kessler2009a}) integrated into the \texttt{pippin} framework \citep{Hinton2020}. We simulate \textbf{size of final sim} SNe in the redshift range $0.05 \leq z_{\mathrm{SN}} \leq 1.3$. The SNe are generated and trained in the same way as for the training of SNN. SALT2 parameters $x_1$ (stretch) and $c$ (colour) are drawn randomly from the intrinsic asymmetric Gaussian distributions described in \citet{Scolnic2016} and "blurred" following the intrinsic scatter model of \citet{Guy2010}, while redshifts are drawn following the volumetric rate evolution of \citet{Frohmaier2019}. The SNe are simulated with explosion epochs $t_0$ uniformly distributed between two months before DES-SN began and two months after it finished in order to account for all SNe that could have been observed by DES-SN. We run mock versions of the DES-SN survey, using the exact cadence, conditions, and zeropoints from the survey itself. All of the detected simulated SNe are passed through the light curve fit of SALT2 \citep{Betoule2014}, as per the implementation in \texttt{SNANA}, and those that fail the light curve cuts outlined in Section \ref{subsubsec:sn_classify} are discarded. We are left with a fraction of the original simulated SNe, and that fraction is dependent on a combination of sky location, explosion epoch, redshift, stretch, and colour. The fraction of recovered SNe (the efficiency) is thus described by a 5-dimensional surface. For the $i$th SN the efficiency $\eta_{\mathrm{SN}, i}$ in field $F$, exploding at time $t_0$ at redshift $z$, with stretch $x_1$ and colour $c$, is:
\begin{equation}
    \eta_{\mathrm{SN},i} (F_i,z_i,t_{0,i},x_1,c) = \left( \frac{N_{\mathrm{obs}}\left(F_i,z_i,t_{0,i},x_1,c\right)}{N_{\mathrm{sim}}\left(F_i,z_i,t_{0,i},x_{1,i},c_i\right)}\right)\,.
\end{equation}

@@

\subsection{Supernova hosts \label{subsec:incompletenss_SN_hosts}}
A further limiting factor in the SN host sample is the requirement of a spectroscopic redshift. The majority of SN host spectroscopic redshifts in DES are provided by the dedicated follow-up survey OzDES, for which the limiting magnitude is around 24 to 24.5 mag in the $r$ band \citep{Lidman2020}. The rate at which a host redshift is successfully measured given an apparent magnitude has been extensively modelled by \citet{Vincenzi2020} who provide the spectroscopic redshift efficiency as a function of host $r$-band magnitude, host galaxy colour, and the year in which the SN was discovered in order to allow for a longer possible spectroscopic exposure time for hosts of SNe discovered earlier in the survey. So that we reduce any bias towards SNe in bright hosts that are easier to obtain spectroscopic redshifts for, the host galaxy spectroscopic redshift efficiency is multiplied by that derived from the SN detection efficiency (Section \ref{subsec:incompleteness_SNe}) to arrive at the final SN efficiency $\eta_{\mathrm{SN},i}$.

>>>>

% Abstract of the paper
\begin{abstract}
We use a sample of 809 photometrically classified type Ia supernovae (SNe Ia) discovered by the Dark Energy Survey (DES) along with 40415 field galaxies to calculate the rate of SNe Ia per galaxy in the redshift range $0.2 < z <0.6$. We recover the positive trend of SN Ia rate as a function of galaxy stellar mass across a broad range of scales $8.5 \leq \log(M_*/\mathrm{M}_{\odot}) \leq 11.25$. We find that the SN Ia rate increases with stellar mass as a power-law with index $0.63 \pm 0.02$, which is consistent with previous work. We use an empirical model of stellar mass assembly to estimate the average star-formation histories (SFHs) of galaxies across the stellar mass range of our measurement. Combining the modelled SFHs with the SN Ia rates to estimate constraints on the SN Ia delay time distribution (DTD), we find the data are fit well by a power-law DTD with index $\beta = -1.14 \pm 0.05$ and normalisation $A = 2.1 \pm0.05 \times 10^{-13}~\mathrm{SNe}~{\mathrm{M}_{\odot}}^{-1}~\mathrm{yr}^{-1}$, which correspond to an overall SN Ia production efficiency $N_{\mathrm{Ia}}/M_* = 0.9~_{-0.7}^{+4.0} \times 10^{-3}~\mathrm{SNe}~\mathrm{M}_{\odot}^{-1}$. Upon splitting the SN sample by properties of the light curves, we find a strong dependence on DTD slope with the SN decline rate, with slower-declining SNe exhibiting a steeper DTD slope. We interpret this as a result of a relationship between intrinsic luminosity and progenitor age, and explore the implications of the result in the context of SN Ia progenitors.
\end{abstract}

@@

Type Ia supernovae (SNe Ia) are explosions of white dwarf stars (WDs). Although SNe Ia show diversity in their observed properties, a large fraction of them ("non-peculiar" SNe Ia) display a small dispersion in their peak brightnesses which can be reduced further through empirical relations between brightness and light curve properties, such as decline rate (stretch) or optical colour \citep{Rust1974,Pskovskii1977,Phillips1993,Tripp1998}. These properties have led SNe Ia to be used extensively by cosmologists to measure relative distances in the Universe \citep{Riess1998,Perlmutter1999}.

@@

The progenitors of a number of core-collapse SNe (CCSNe) have been identified in high-resolution pre-explosion images \citep{Smartt2009,Eldridge2013}, allowing detailed analysis of the latter stages of massive star evolution. However, this approach has not yet been successful in identifying a progenitor of a SN Ia (e.g. \citealt{Graur2014a,Kelly2014,Graur2019}, although see \citet{McCully2014} for evidence of a progenitor for a type Iax SN). Conversely, it has been possible to search for surviving remnants of the binary system -- main sequence stars or WDs left behind at the SN location or kicked out at high velocity \citep[e.g.][]{Schaefer2012,Ruiz-Lapuente2018,Kerzendorf2018,Kerzendorf2019}. Promising recent results include that from \citet{Shen2018} who discovered three high-velocity WDs consistent with models of double-degenerate, double-detonation SNe Ia. However, without an unambiguous observation of a SN Ia progenitor system or remnant companion, there is as yet no direct evidence ruling out any progenitor channel from providing a significant contribution to the total population of SNe Ia.

Another method of inferring progenitor channels is the study of the SN Ia delay time distribution (DTD). The DTD describes the rate at which SNe Ia occur as a function of the delay time $\tau$ since an episode of star-formation, and thus carries characteristic signatures of the progenitor channel (see \citealt{Wang2012} for a review of the theory, and \citealt{Maoz2017} for a review of observations). The SD scenario produces a broad range of functional forms for the DTD, most of which fail to account for long delay times \citep{Graur2014}. On the other hand, most variants of the DD scenario predict a power-law of the form $\tau^{\beta}$, with $\beta \sim -1$ \citep[e.g][]{Ruiter2009,Mennekens2010}.

@@

\citet[][hereafter C14]{Childress2014} showed that the $A+B$ approximation arises as a direct consequence of the combination of a power-law DTD with the average SFHs of galaxies -- the prompt component proportional to the amount of on-going star-formation at the epoch of observation, and the tardy component caused by the fact that massive galaxies experienced high SFRs several Gyrs ago. Recent advances in integral field spectroscopy (IFS) allow for an extension of the SFHR method, by reconstructing the SFH for hundreds or thousands of local regions of each SN Ia host galaxy. Using this method, \citet{Castrillo2020} find a power-law slope of $-1.1\pm0.3$, while \citet{Chen2021} find $-1.4\pm0.3$.

The majority of observational evidence based on studies of SN Ia populations thus points towards a DD scenario for most, if not all, SNe Ia. However, finding self-consistent progenitor and explosion models that recreate the observed luminosity function as well as correlations between luminosity, light curve parameters, and host galaxy properties has proven difficult. In particular, simulations based around explosions of $M_{\mathrm{Ch}}$ WDs (linked strongly with the SD scenario but also with many DD scenarios) find difficulty in reproducing the light curves of "normal" SNe Ia as well as "peculiar" objects (\citealt{Ropke2007,Sim2013,Blondin2017}; see \citealt{Maoz2014,Jha2019} for overviews). In recent years, attention directed towards explosions of sub-$M_{\mathrm{Ch}}$ WDs triggered by double detonations (primarily related to a DD scenario) has led to promising results \citep[e.g.][]{Shen2017,Shen2018,Townsley2019,Gronow2020,Shen2021} although they still struggle to match observations at late times in the light curve evolution \citep{Shen2021,Gronow2021}. An additional factor in support of the sub-$M_{\mathrm{Ch}}$ model is that the SN luminosity is related to the mass of the primary WD, which itself is likely to be related to its age (although this relation is probably complicated by other factors such as accretion rate, metallicity, and the composition of the companion), thereby providing an explanation for observed relation between light curve stretch and stellar age \citep{Rigault2013,Rigault2018,Rose2019,Nicolas2020}. With many models showing promising similarities to observations but each subject to its own drawbacks, it is becoming accepted that more than one progenitor scenario may contribute significantly to the overall population of "normal" SNe Ia; detailed observations are thus required in order to place constraints on the relative fractions of each possible progenitor channel.

@@

In Section \ref{sec:data} we introduce our large sample of SNe Ia from the Dark Energy Survey (DES) as well as our deep sample of field galaxies that provide an effectively complete control sample from which to measure the rate. We describe our detailed handling of SN and galaxy incompleteness in Section \ref{sec:incompleteness}. We present the SN Ia rate per galaxy in Section \ref{sec:rates} and show that it is consistent with previous works. We outline our modelling of star-formation histories and our novel constraints on the DTD parameters in Section \ref{sec:model}. In Section \ref{sec:split_x1_c} we investigate how the DTD differs among sub-populations of SNe with different light curve characteristics, in particular the stretch, and show stretch to be strongly dependent on progenitor age. We conclude in Section \ref{sec:conclusion} by discussing the implications of the results.
Where relevant, we assume a spatially flat $\Lambda$CDM cosmology with $\Omega_m = 0.3$ and $H_0 = 70 \mathrm{~km~s}^{-1}~\mathrm{Mpc}^{-1}$. Unless otherwise stated, we assume Gaussian measurement uncertainties quoted at the $1\sigma$ level, and we quote the posterior median and $1\sigma$ credible intervals on derived parameters. Magnitudes are quoted in the AB system \citep{Oke1983}.

\begin{table}
	\centering
	\caption{Numbers of field galaxies and SN hosts passing various quality cuts. SNe are derived from must already have passed the masked chips and good redshift cuts before reaching this stage.}
	\label{tab:cuts}
	\begin{tabular}{lccr} % four columns, alignment for each
		\hline
		Cut & N (field galaxies)  & N (SN hosts)\\
		\hline
		Chosen fields & 1364311  & 1441\\
	    Kron mag in all bands & 1069004  & 1439 \\
	    Masked chips & 816950  & - \\
	    Has ugrizJHK photometry & 545748  & -\\
	    Edge of chip & 481731 & 1401 \\
	    Star/galaxy separation & 400051 &  1259\\
	    Has good redshift & 395034 & - \\
	    SNR $\geq 3$& 338256  & 1259 \\
	    $m_r \leq 24.5$ & 196109 &  1254 \\
	    $0.2 \leq z \leq 0.6$ & 48177 & 809\\

@@

%In an ideal rate measurement a volume-limited and complete sample of galaxies would be observed for a given time, and the number of SNe observed in each galaxy simply counted. In practice, apart from at very low redshift where the absolute numbers do not support robust statistical analyses, this is not possible: SNe are detected, classified and followed up based on complicated selection criteria; galaxy surveys are incomplete at faint magnitudes; the timescale of observing seasons cuts off SNe at the beginning and end of each season.
In order to measure the rate of SNe Ia per galaxy per year, we require a sample of SNe Ia, as well as a sample of all of the possible galaxies ("field" galaxies) that those SNe could have exploded in. In practice, SN and galaxy surveys do not cover the same sky area, redshift ranges, times, and have different selection biases. In this section, we introduce our SN and field galaxy samples and in Section \ref{sec:incompleteness} we describe our method of correcting the selection effects for both the SN and field galaxy samples.

\label{sec:data}
\subsection{Dark Energy Survey supernova programme \label{subsec:des}}
To derive our sample of SNe and field galaxies, we make use of the Dark Energy Survey (DES). The DES Supernova Programme (DES-SN) was a transient survey based on five six-month seasons of observations of ten southern hemisphere fields with the Dark Energy Camera (DECam; \citealt{Flaugher2015}), designed primarily to measure the light curves of SNe Ia for use as cosmological distance indicators. Transients were detected and processed using a difference imaging program \citep{Kessler2015} and vetted using a machine learning (ML) algorithm \citep{Goldstein2015}. Spectral follow-up of live SN candidates was performed on a suite of large optical telescopes \citep{Smith2020a}, leading to the initial publication of a measurement of cosmological parameters using 207 spectroscopically confirmed SNe Ia (\citealt{DESCollaboration2018a}, and references therein). Spectroscopic redshifts for host galaxies come from the OzDES Global Redshift Catalog (GRC; \footnote{\url{https://docs.datacentral.org.au/ozdes/overview/dr2/}} \citealt{Yuan2015,Childress2017,Lidman2020}), which comprises galaxies targetted as DES-SN hosts with the OzDES programme as well as redshifts from legacy catalogues in the DES-SN fields.
\subsection{Supernovae \label{subsec:host_sample}}
\subsubsection{Photometric classification and quality cuts \label{subsubsec:sn_classify}}
With over 30,000 transients passing ML selection it was not possible to obtain spectral follow up of every object. Instead we make use of photometric classification in the form of the recurrent neural network classifier \texttt{superNNova} (SNN; \citealt{Moller2019}), foloowing the approach of \citet{Scolnic2020}. A full description of the training and fitting of SNN will be presented in Vincenzi et al. \textit{in prep} and M\"{o}ller et al. \textit{in prep}; we provide a brief overview here.

@@

Since the vast majority of objects detected in the SN Deep catalogue do not have a spectroscopic redshift measurement, we rely on photometric redshifts (photo-$z$s).
Photo-$z$s are taken, from the DES Y3A2\_DEEP catalogue of \citet{Hartley2020}. Y3A2\_DEEP makes use of the same deep DES optical photometry that was used in SN Deep for a subset of the fields (SN-X3, SN-C3 and SN-E2), but adds DECam $u$-band data and near-infrared (NIR) $J$, $H$, and $K$-band imaging from the ultraVISTA and VIDEO surveys. The deep optical photometry in Y3A2\_DEEP was stacked using a different technique to that of SN Deep, but the resulting photometry is consistent within $1\sigma$ uncertainties (\citealt{Wiseman2020},Meledorf et al. \textit{in prep}). \citet{Hartley2020} estimated photo-$z$s using the \texttt{eaZy} code \citep{Brammer2008}. Photo-$z$ accuracy from \citet{Hartley2020} at $17 < i<24$ is estimated around 0.03 as quantified by the Normalised Absolute Median Deviation (NMAD). This accuracy is degraded to 0.07 at $17 < i<26$, a range that includes all galaxies in our sample, but is more than adequate for the purposes of this work, where all field galaxies will be grouped into stellar mass bins of $\log \left(\Delta M_*\right) = 0.25$.

\begin{figure}
    \centering
    \includegraphics[width=.5\textwidth]{figs/mass_hist_linlog_sn_field_BC03.png}
    \caption{Upper: Distribution of the stellar mass of the SN host (green dot-dashed) and field galaxies (orange and grey solid) used in the analysis. Lower: As upper for field galaxies, but shown on a log scale.  The distribution closely matches data from the ZFOURGE survey at $0.5<z<0.75$ (magenta dots). The difference between the raw counts and the $V_{\mathrm{max}}$ corrected counts (Section \ref{sec:incompleteness}) is minimal. 
    \label{fig:mass_hist_hosts_field}}
\end{figure}

@@

    \item field galaxies are limited to unmasked region of the SN-X3 field \citep{Hartley2020};

@@

    \item to minimise stellar contamination, objects must have a value of $<0.95$ in at least one band for the star/galaxy (S/G) separation metric \texttt{CLASS\_STAR} provided by \texttt{Source Extractor} \citep{Bertin1996};

@@

    \item objects must have a spectroscopic redshift measurement or a photometric redshift estimate with a well-defined peak in redshift  probability density;

@@

The numbers of SN hosts and field galaxies passing these cuts are listed in Table \ref{tab:cuts}. The final samples comprise 809 SNe and their host galaxies and 40,415 field galaxies. The volume-weighted mean redshifts are 0.50 for both SNe and field galaxies.

@@

We estimate global galaxy properties for both the SN host and field galaxies that pass the quality cuts by fitting the photometry with stellar population templates in a method outlined by \citet{Sullivan2006} and consistent to that used in previous DES-SN analyses \citep{Smith2020,Wiseman2020,Kelsey2021}. For SN hosts the redshift is fixed at the spectroscopically determined value, while for field galaxies we fix it at either the spectroscopic value if one exists in the OzDES GRC, or more commonly the photometrically derived value. We use the stellar population templates of \citet{Bruzual2003} and adopt a \citet{Chabrier2003} initial mass function (IMF). The fitting procedure returns a best fitting template and corresponding stellar mass ($M_*$). Upper and lower bounds on stellar mass estimates are taken as the extreme values that correspond to templates that are consistent with the data (given the photometric uncertainties) according to their $\chi^2$ statistic, as per \citet{Sullivan2006}. To check that the choice of templates does not bias the result, we also fit the galaxies using the P\'EGASE.2 templates \citep{Fioc1997,LeBorgne2002} and a \citet{Kroupa2001} IMF, and find results consistent within measurement uncertainties. This is consistent with the findings of \citet{Smith2020}.

@@

\section{Incompleteness corrections}
\label{sec:incompleteness}
The simple ratio of the number of SN hosts and field galaxies presented at the end of the previous section provides a first approximation of the SN rate per galaxy, and the addition of a factor equal to the survey duration normalises the rate to per year. However, both the SN and field galaxy samples introduced in Sections \ref{subsec:host_sample} and \ref{subsec:field_sample} are affected by incompleteness, which is likely to be the dominant systematic effect in the analysis. In this section we describe our method of correcting for various sources of incompleteness in the data.

@@

To correct for these incompleteness, we follow a similar method to that used in the Palomar Transient Factory rates analyses of \citet{Frohmaier2019} and \citet{Frohmaier2020}. We run a suite of simulations using an identical method to that described in detail in \citet{Vincenzi2020}. We simulate SNe using the SuperNova ANAlysis software (\texttt{SNANA}; \citealt{Kessler2009a}) integrated into the \texttt{pippin} framework \citep{Hinton2020}. We simulate $1.1\times 10^6$ SNe in the redshift range $0.05 \leq z_{\mathrm{SN}} \leq 1.3$. The SNe are generated and trained in the same way as for the training of SNN. SALT2 parameters $x_1$ (stretch) and $c$ (colour) are drawn randomly from the intrinsic asymmetric Gaussian distributions described in \citet{Scolnic2016} and "blurred" following the intrinsic scatter model of \citet{Guy2010}, while redshifts are drawn following the volumetric rate evolution of \citet{Frohmaier2019}. The SNe are simulated with explosion epochs $t_0$ uniformly distributed between two months before DES-SN began and two months after it finished in order to account for all SNe that could have been observed by DES-SN. We run mock versions of the DES-SN survey, using the exact cadence, conditions, and zeropoints from the survey itself. All of the detected simulated SNe are passed through the light curve fit of SALT2 \citep{Betoule2014}, as per the implementation in \texttt{SNANA}, and those that fail the light curve cuts outlined in Section \ref{subsubsec:sn_classify} are discarded. We are left with a fraction of the original simulated SNe, and that fraction is dependent on a combination of sky location, explosion epoch, redshift, stretch, and colour. The fraction of recovered SNe (the efficiency) is thus described by a 5-dimensional surface. For the $i$th SN the efficiency $\eta_{\mathrm{SN}, i}$ in field $F$, exploding at time $t_0$ at redshift $z$, with stretch $x_1$ and colour $c$, is:
\begin{equation}
    \eta_{\mathrm{SN},i} (F_i,z_i,t_{0,i},x_1,c) = \left( \frac{N_{\mathrm{obs}}\left(F_i,z_i,t_{0,i},x_1,c\right)}{N_{\mathrm{sim}}\left(F_i,z_i,t_{0,i},x_{1,i},c_i\right)}\right)\,.
\end{equation}

@@

\subsection{Supernova hosts \label{subsec:incompletenss_SN_hosts}}
A further limiting factor in the SN host sample is the requirement of a spectroscopic redshift. The majority of SN host spectroscopic redshifts in DES are provided by the dedicated follow-up survey OzDES, for which the limiting magnitude is around 24 to 24.5 mag in the $r$ band \citep{Lidman2020}. The rate at which a host redshift is successfully measured given an apparent magnitude has been extensively modelled by \citet{Vincenzi2020} who provide the spectroscopic redshift efficiency as a function of host $r$-band magnitude, host galaxy colour, and the year in which the SN was discovered in order to allow for a longer possible spectroscopic exposure time for hosts of SNe discovered earlier in the survey. So that we reduce any bias towards SNe in bright hosts that are easier to obtain spectroscopic redshifts for, the host galaxy spectroscopic redshift efficiency is multiplied by that derived from the SN detection efficiency (Section \ref{subsec:incompleteness_SNe}) to arrive at the final efficiency for the $i$th SN, $\eta_{\mathrm{SN},i}$.
